Project Equalizer

### 2007-03-13 WhitePapers/ProjectEqualizer/paper.tex
[ ] May break build [ ] Breaks existing applications [ ] Bugfix [ ] New Feature [ ] Cleanup [ ] Optimization [x] Documentation
Description: Update use cases web page

Today we have satisfying, though often proprietary, solutions for the
transparent approach\footnote{e.g., Chromium from Stanford University,
VGP from ModViz, SVN from IBM}. Distributed scene graphs are beginning
to emerge\footnote{e.g., OpenSG, ScaleViz from Mercury Computer Systems} as
well. There is currently no general programming framework to create
parallel graphics applications. Furthermore, all existing solutions are
'islands', that is, they are configured separately and do possibly
interfere with each other when running simultanously on the same
system. Equalizer addresses these issues.

Unlike the HPC community, visualization software developers do not
widely address performance bottlenecks by parallelizing
their applications. One of the reasons for the lack of parallel graphics
applications is the lack of a standard programming interface, which
implements specific knowledge needed to create scalable visualization
applications. Equalizer provides a programming framework for high
performance visualization.

@@

Equalizer is a project to create new software to facilitate the
development and usage of software for multipipe
graphics systems by addressing the shortcomings of the current
software situation. It adds the two missing components, a
parallel programming framework and a resource management system, and
integrates existing solutions, such as a transparent layer, scene graphs
and remote visualization capabilities.

Equalizer addresses all visualization systems. An Equalizer application
can run unmodified on a laptop, a multipipe workstation, a graphics
cluster or a multipipe shared memory visualization server.

@@

\begin{description}
\item[Lower Development Costs] The common functionality of any multipipe
  application is outsourced to the Equalizer framework, thus freeing the
  software developers to reimplement basic algorithms. The expertise for
  high performance visualization is partly embedded into Equalizer, and
  has not to be developed inhouse. The open development approach ensures
  that Equalizer delivers the functionality most needed by the community.
\item[Higher Functionality] Additional features, which would not be
  cost-effective to implement in the application, are provided by the
  Equalizer framework. New features will become available over time,
  and will often be usable by the application with no or minimal code
  changes.
\item[Ease of Use] The central resource management of Equalizer requires
  only a one-time setup, thus minimizing the impact for end users. Once
  a visualization system is set up, unmodified and scalable OpenGL
  applications can be run without any modification. Simple
  configurations, especially shared memory systems, will be configured
  automatically.
\item[Reduced Development Risk] The open source license of Equalizer is
  the guarantee for software developers that the product will be
  available 'forever'. Critical features or bugs, in the event that they are
  not addressed by the maintainers, can still be addressed internally.
\end{description}

@@

For standalone usage on simple machines, the Equalizer client library
will be able to automatically configure and use a builtin server, to
facilitate the deployment on workstations. Naturally, certain
optimisations, such as the per system loadbalancing is lost without the
use of a dedicated server.

\subsection{Integration}
The following components will be based on existing open source software, which
will be integrated with Equalizer. This integration enables a seamless
execution and end user experience. On one hand, by relying on the
Equalizer server for configuration, an visualization systems has only
one set of configuration files, instead of disjunct configurations for
each of the tools. On the other hand, the central resource management improves
the performance and interoperability between applications running on
the same system.

@@

\subsubsection{Scene Graphs}
The integration with scene graphs enables existing applications
using such a scene graph to quickly be able to take advantadge of
multipipe systems.

@@

Currently, the programming framework is usable for the development of
commercial multipipe applications by an experienced OpenGL
programmer. Version 0.2 added support for scalable rendering to show the
benefits of scalable rendering on commodity graphics clusters and shared
memory systems.

>>>>

Today we have satisfying, though often proprietary, solutions for the
transparent approach\footnote{Chromium from Tungsten Graphcis/Stanford
  University, VGP from ModViz, SVN from IBM}. Distributed scene graphs
are beginning to emerge\footnote{e.g., OpenSG, ScaleViz from Mercury
  Computer Systems} as well. There is currently no general programming
framework to create parallel graphics applications. Furthermore, all
existing solutions are 'islands', that is, they are configured
separately and do possibly interfere with each other when running
simultanously on the same system. The goal of Equalizer is to address
these issues.

Unlike the HPC community, visualization software developers do not yet
widely address performance bottlenecks by parallelizing their
applications. One of the reasons for the lack of parallel graphics
applications is the lack of a standard programming interface, which
implements specific knowledge needed to create scalable visualization
applications. Equalizer provides such a programming framework for high
performance visualization.

@@

Equalizer is a project to create new software to facilitate the
development and usage of software for multipipe graphics systems by
addressing the shortcomings of the current software situation. It adds
the two missing components, a parallel programming framework and a
resource management system, and will integrate existing solutions, such
as a transparent layer, scene graphs and remote visualization
capabilities.

Equalizer addresses all types of visualization systems. An Equalizer
application can run unmodified on a laptop, a multipipe workstation, a
graphics cluster or a multipipe shared memory visualization server.

@@

\begin{description}
\item[Lower Development Costs] The common functionality of any multipipe
  application is outsourced to the Equalizer framework, thus freeing the
  software developers to reimplement basic parallel rendering
  algorithms. The expertise for high performance visualization is partly
  embedded into Equalizer, and has not to be developed in house. The
  open development approach ensures that Equalizer delivers the
  functionality most needed by the community.
\item[Higher Functionality] Additional features, which would not be
  cost-effective to implement in the application, are provided by the
  Equalizer framework. New features will become available over time, and
  will often be usable by the application with no or minimal code
  changes.
\item[Ease of Use] The central resource management of Equalizer requires
  only a one-time setup, thus minimizing the impact for end users. Once
  a visualization system is set up, unmodified and scalable OpenGL
  applications can be run without any modification. Simple
  configurations, especially shared memory systems, will be configured
  automatically.
\item[Reduced Development Risk] The open source license of Equalizer is
  the guarantee for software developers that the product will be
  available 'forever'. Critical features or bugs, in the event that they
  are not addressed by the maintainers, can still be addressed
  internally.
\end{description}

@@

For standalone usage on simple machines, the Equalizer client library
will be able to automatically configure and use a built-in server, to
facilitate the deployment on workstations. Naturally, certain
optimisations, such as the per-system loadbalancing is lost without the
use of a dedicated, system-wide server.

\subsection{Integration}
The following components will be based on existing open source software, which
will be integrated with Equalizer. This integration enables a seamless
execution and end user experience. On one hand, by relying on the
Equalizer server for configuration, a visualization systems has only
one set of configuration files, instead of disjunct configurations for
each of the tools. On the other hand, the central resource management improves
the performance and interoperability between applications running on
the same system.

@@

\subsubsection{Scene Graphs}
The integration with scene graphs enables existing applications using
such a scene graph to quickly be able to take advantadge of parallel
rendering on multipipe systems.

@@

Currently, the programming framework is usable for the development of
commercial multipipe applications by an experienced OpenGL
programmer. Version 0.2 added support for scalable rendering, showing
the benefits of scalable rendering on commodity graphics clusters and
shared memory systems.
